---
title: "Midterm"
author: "Emmy Niswonger"
date: "`r Sys.Date()`"
output: html_document
---
Midterm Assignment

Use the material you have learned this class to build a dashboard studying the mathematics and/or reading proficiency scores for 2020-2021.  You have flexibility to specify the story in your dashboard. Be creative!!!

Load the csv files for the math achievement and reading achievement scores
```{r}
library(dplyr)
library(tidyverse)
options(dplyr.summarise.inform = FALSE)

rla_achievement <- read_csv("rla-achievement-lea-sy2020-21.csv")
math_achievement <- read_csv("math-achievement-lea-sy2020-21 (1).csv")
```

##Reading Proficiency Scores

Filter the reading scores data for all of the students in the school statewide with at least 1000 valid entries

Now, mutate the PCTPROF into a numeric factor and find the mean percentage per state
```{r}
filtered_data_reading_scores <- rla_achievement %>% 
  filter(CATEGORY == "ALL", GRADE == "00", NUMVALID >= 1000) %>%
  group_by(State = STNAM,
           County = LEANM) %>%
  mutate(PCTPROF = as.numeric(PCTPROF, na.rm=TRUE)) %>%
summarise(Percent_Reading_Proficiency = mean(PCTPROF, na.rm=TRUE))
View(filtered_data_reading_scores)
```

Find the three highest mean reading proficiency percentages and the three lowest mean proficiency percentages 
```{r}
highest_percentages_reading <- filtered_data_reading_scores %>%
  slice_max(Percent_Reading_Proficiency, n=3)

lowest_percentages_reading <- filtered_data_reading_scores %>%
  slice_min(Percent_Reading_Proficiency, n=3)
```

Now, merge the mean highest and lowest reading proficiency percentages into a dataframe 
```{r}
average_proficiency_reading <- rbind(highest_percentages_reading, lowest_percentages_reading)
```

##Math Proficiency Scores

Filter the math scores data for all of the students in the school statewide with at least 1000 valid entries
```{r}
filtered_data_math_scores <- math_achievement %>% 
  filter(CATEGORY == "ALL", GRADE == "00", NUMVALID >= 1000) %>%
  group_by(State = STNAM,
           County = LEANM) %>%
  mutate(PCTPROF = as.numeric(PCTPROF)) %>%
summarise(Percent_Math_Proficiency = mean(PCTPROF, na.rm = TRUE))
View(filtered_data_math_scores)
```

Find the three highest mean math proficiency percentages and the three lowest mean proficiency percentages 
```{r}
highest_percentages_math <- filtered_data_math_scores %>%
  slice_max(Percent_Math_Proficiency, n=3)

lowest_percentages_math <- filtered_data_math_scores %>%
  slice_min(Percent_Math_Proficiency, n=3)
```

Now, merge the mean highest and lowest reading proficiency percentages into a dataframe 
```{r}
average_proficiency_math <- rbind(highest_percentages_math, lowest_percentages_math)
```

## Output for both reading and math average proficiency scores 

```{r}
average_scores <- cbind(average_proficiency_math, average_proficiency_reading)
```
